---
title: "Suitability"
author: "AnnaClaire Marley"
date: "5/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#load packages
library(tidyverse)
library(raster)
library(sdmpredictors)
library(maps)
library(mapdata)
library(rnaturalearth)
```


# Suitability 

**General Seaweed:**
- Depth: 10-200m
- N:P ratio: between 4 and 80
-----------------------------------------
- Sea surface temperature range: 0-35°C
- Current velocity: 0.1-1 m/s

*Sargassum* specific:
- Optimal sea surface temp: 18-30°C
- Optimal salinity: 24-42 psu

*Euchema* specific:
- Optimal sea surface temp: 25-30°C
- Optimal salinity: 28-34 psu

*Gracilaria* specific:
- Optimal sea surface temp: 15-30°C
- Optimal salinity: 15-25 psu

*Gracilaria spp* from Lisa specific:
- Optimal sea surface temp: 20-30°C
- Optimal salinity: 25-40 psu
- Current velocity: 0.05-0.5 m/s

**Seaweed Parameters**
```{r}
# depth
min_depth <- -10
max_depth <- -200

# Nitrate:Phosphate ratio
n_p_ratio_min <- 4
n_p_ratio_max <- 80

# sea surface temperature
min_sst_value <- 18
max_sst_value <- 30

# salinity
min_sal_value <- 24
max_sal_value <- 42

# current velocity
min_curvel_value <- 0.1
max_curvel_value <- 1

```

```{r}
# set up path to load activate layers
path <- "../data/Active_Data_Layers/"
```


Depth
```{r}

# read in raster
depth_mask <- raster(paste0(path,"/depth_mask_bo.tif"))

# Reclassification matrix 
rcl_matrix_depth <- c(-Inf, max_depth, 0,
                   max_depth, min_depth, 1,
                   min_depth, 0, 0,
                   0, Inf, 0)

# Reclassify 
depth_binary <- reclassify(depth_mask, rcl = rcl_matrix_depth)

# Convert to data frame
depth_df <- raster::rasterToPoints(depth_binary) %>%
  as.data.frame()%>% 
  mutate(suit = case_when(
    depth_mask_bo == 1 ~ "suitable",
    depth_mask_bo == 0 ~ "unsuitable"
  ))


# Plot reclassified layer
# plot(depth_binary)

world <- ne_countries(scale = "medium", returnclass = "sf")

depth_map <-
  ggplot() +
  geom_raster(data = depth_df, aes(x = x, y = y, fill = suit)) +
  geom_sf(data = world, fill = "cornsilk3") +
  coord_sf(xlim = c(-100, -80), ylim = c(22, 32)) +
  scale_fill_manual(values = c("olivedrab", "lightgray")) +
  theme_minimal() +
  labs(fill = "",
       x = "longitude",
       y = "latitude", 
       title = "Depth Suitability") +
  theme(legend.background = element_rect(fill = "#3e5c8b", color = "#3e5c8b"),
        legend.title = element_text(colour = "#e0bf53"),
        legend.text = element_text(colour = "#fffff0"),
        plot.background = element_rect(fill = "#3e5c8b", color = "#3e5c8b"),
        axis.text = element_text(colour = "#fffff0"), 
        axis.title = element_text(colour = "#fffff0"),
        plot.title = element_text(colour = "#fffff0", hjust = 0.5),
        panel.grid.major.x = element_line(colour = "#fffff0"),
        panel.grid.major.y = element_line(colour = "fffff0"))

depth_map 


#ggsave("depth_formatted.png", plot = depth_map)

#Use the magick package to trim off the white edges.
library(magick)
depth_map_trim <- magick::image_read("depth_formatted.png") %>% magick::image_trim()


#image_write(depth_map_trim, path = "depth_map_trim.png", format = "png")

```

```{r}
# load function to relcassify rasters
source('../functions/reclass_data.R')
```

Max SST
```{r}
# read in raster
max_sst <- raster(paste0(path,"/max_sst_mask.tif"))

# reclassify
max_sst_binary <- reclass_data(raster = max_sst, max_thresh = max_sst_value, 
                        one_thresh = TRUE, which_thresh = "MAX")

plot(max_sst)

# plot reclassified layer
plot(max_sst_binary)

```


Min SST
```{r}
# read in raster
min_sst <- raster(paste0(path,"/min_sst_mask.tif"))

plot(min_sst)

# reclassify
min_sst_binary <- reclass_data(raster = min_sst, min_thresh = min_sst_value, 
                        one_thresh = TRUE, which_thresh = "MIN")


# plot reclassified layer
plot(min_sst_binary)

```

N:P ratio
```{r}
# read in raster
n_p_ratio <- raster(paste0(path,"/n_p_ratio.tif"))

# reclassify
np_binary <- reclass_data(raster = n_p_ratio, min_thresh = n_p_ratio_min,
                          max_thresh = n_p_ratio_max, one_thresh = FALSE)

# plot reclassified layer
# plot(np_binary)

# Convert to data frame
np_df <- raster::rasterToPoints(np_binary) %>%
  as.data.frame() %>% 
  mutate(suit = case_when(
    n_p_ratio == 1 ~ "suitable",
    n_p_ratio == 0 ~ "unsuitable"
  ))


# Plot reclassified layer
# plot(depth_binary)

np_map <-
  ggplot() +
  geom_raster(data = np_df, aes(x = x, y = y, fill = suit)) +
  geom_sf(data = world, fill = "cornsilk3") +
  coord_sf(xlim = c(-100, -80), ylim = c(22, 32)) +
  scale_fill_manual(values = c("olivedrab", "lightgray")) +
  theme_minimal() +
  labs(fill = "",
       x = "longitude",
       y = "latitude", 
       title = "Nitrogen and Phosphorus Suitability") +
  theme(legend.background = element_rect(fill = "#3e5c8b", color = "#3e5c8b"),
        legend.title = element_text(colour = "#e0bf53"),
        legend.text = element_text(colour = "#fffff0"),
        plot.background = element_rect(fill = "#3e5c8b", color = "#3e5c8b"),
        axis.text = element_text(colour = "#fffff0"), 
        axis.title = element_text(colour = "#fffff0"),
        plot.title = element_text(colour = "#fffff0", hjust = 0.5),
        panel.grid.major.x = element_line(colour = "#fffff0"),
        panel.grid.major.y = element_line(colour = "fffff0"))

np_map 


#ggsave("np_ratio_formatted.png", plot = np_map)

#Use the magick package to trim off the white edges.
library(magick)
np_ratio_map_trim <- magick::image_read("np_ratio_formatted.png") %>% magick::image_trim()


#image_write(np_ratio_map_trim, path = "np_ratio_map_trim.png", format = "png")
```

Salinity
```{r}
# read in files
sal_max <- raster(paste0(path,"/sal_max_mask.tif"))
sal_min <- raster(paste0(path,"/sal_min_mask.tif"))


# reclassify
sal_max_binary <- reclass_data(raster = sal_max, max_thresh = max_sal_value, 
                        one_thresh = TRUE, which_thresh = "MAX")
sal_min_binary <- reclass_data(raster = sal_min, min_thresh = min_sal_value, 
                        one_thresh = TRUE, which_thresh = "MIN")

# plot
# plot(sal_max_binary)
# plot(sal_min_binary) # use this one for total salinity suitability

# Convert to data frame
salinity_df <- raster::rasterToPoints(sal_min_binary) %>%
  as.data.frame() %>% 
  mutate(suit = case_when(
    sal_min_mask == 1 ~ "suitable",
    sal_min_mask == 0 ~ "unsuitable"
  ))


salinity_map <-
  ggplot() +
  geom_raster(data = salinity_df, aes(x = x, y = y, fill = suit)) +
  geom_sf(data = world, fill = "cornsilk3") +
  coord_sf(xlim = c(-100, -80), ylim = c(22, 32)) +
  scale_fill_manual(values = c("olivedrab", "lightgray")) +
  theme_minimal() +
  labs(fill = "",
       x = "longitude",
       y = "latitude", 
       title = "Salinity Suitability") +
   theme(legend.background = element_rect(fill = "#3e5c8b", color = "#3e5c8b"),
        legend.title = element_text(colour = "#e0bf53"),
        legend.text = element_text(colour = "#fffff0"),
        plot.background = element_rect(fill = "#3e5c8b", color = "#3e5c8b"),
        axis.text = element_text(colour = "#fffff0"), 
        axis.title = element_text(colour = "#fffff0"),
        plot.title = element_text(colour = "#fffff0", hjust = 0.5),
        panel.grid.major.x = element_line(colour = "#fffff0"),
        panel.grid.major.y = element_line(colour = "fffff0"))

salinity_map 

#ggsave("salinity_formatted.png", plot = salinity_map)

#Use the magick package to trim off the white edges.
library(magick)
salinity_map_trim <- magick::image_read("salinity_formatted.png") %>% magick::image_trim()


#image_write(salinity_map_trim, path = "salinity_map_trim.png", format = "png")

```

Current Velocity

```{r}
# read in raster
curvel_max <- raster(paste0(path,"/curvel_max_mask.tif"))
curvel_min <- raster(paste0(path,"/curvel_min_mask.tif"))

# reclassify
curvel_max_binary <- reclass_data(raster = curvel_max, max_thresh = max_curvel_value, 
                        one_thresh = TRUE, which_thresh = "MAX")
curvel_min_binary <- reclass_data(raster = curvel_min, min_thresh = min_curvel_value, 
                        one_thresh = TRUE, which_thresh = "MIN")

######### we need to figure out how to combine rasters and clip overlap, but for now just using max velocity variable
# plot
#plot(curvel_max_binary) # selecting max velocity for now...
#plot(curvel_min_binary)

# Convert to data frame
velocity_df <- raster::rasterToPoints(curvel_max_binary) %>%
  as.data.frame() %>% 
  mutate(suit = case_when(
    curvel_max_mask == 1 ~ "suitable",
    curvel_max_mask == 0 ~ "unsuitable"
  ))


velocity_map <-
  ggplot() +
  geom_raster(data = velocity_df, aes(x = x, y = y, fill = suit)) +
  geom_sf(data = world, fill = "cornsilk3") +
  coord_sf(xlim = c(-100, -80), ylim = c(22, 32)) +
  scale_fill_manual(values = c("olivedrab", "lightgray")) +
  theme_minimal() +
  labs(fill = "",
       x = "longitude",
       y = "latitude", 
       title = "Current Velocity Suitability") +
  theme(legend.background = element_rect(fill = "#3e5c8b", color = "#3e5c8b"),
        legend.title = element_text(colour = "#e0bf53"),
        legend.text = element_text(colour = "#fffff0"),
        plot.background = element_rect(fill = "#3e5c8b", color = "#3e5c8b"),
        axis.text = element_text(colour = "#fffff0"), 
        axis.title = element_text(colour = "#fffff0"),
        plot.title = element_text(colour = "#fffff0", hjust = 0.5),
        panel.grid.major.x = element_line(colour = "#fffff0"),
        panel.grid.major.y = element_line(colour = "fffff0"))

velocity_map 

#ggsave("current_velocity_formatted.png", plot = velocity_map)

#Use the magick package to trim off the white edges.
library(magick)
current_velocity_map_trim <- magick::image_read("current_velocity_formatted.png") %>% magick::image_trim()


#image_write(current_velocity_map_trim, path = "current_velocity_map_trim.png", format = "png")

```


### Final Maps

```{r}
# source functions to save maps created
source("../functions/save_map.R")
```

Suitability areas -- all bio-oracle
```{r}

suitable_areas <- raster::overlay(np_binary,
                                  depth_binary, 
                                  #max_sst_binary,
                                  #min_sst_binary, 
                                  #sal_max_binary, 
                                  sal_min_binary, 
                                  curvel_max_binary, 
                                  #curvel_min_binary,
                          fun = function(a, b, c, d){a*b*c*d})

# plot
#plot(suitable_zoom)
#map('world',fill=T,add=T,col='gray')


# Convert to data frame
suitable_areas_df <- raster::rasterToPoints(suitable_areas) %>%
  as.data.frame() %>% 
  mutate(suit = case_when(
    layer == 1 ~ "suitable",
    layer == 0 ~ "unsuitable"
  ))


suitable_areas_map <-
  ggplot() +
  geom_raster(data = suitable_areas_df, aes(x = x, y = y, fill = suit)) +
  geom_sf(data = world, fill = "cornsilk3") +
  coord_sf(xlim = c(-100, -80), ylim = c(22, 32)) +
  scale_fill_manual(values = c("olivedrab", "lightgray")) +
  theme_minimal() +
  labs(fill = "",
       x = "longitude",
       y = "latitude", 
       title = "Suitable Seaweed Aquaculture Areas") +
  theme(legend.background = element_rect(fill = "#3e5c8b", color = "#3e5c8b"),
        legend.title = element_text(colour = "#e0bf53"),
        legend.text = element_text(colour = "#fffff0"),
        plot.background = element_rect(fill = "#3e5c8b", color = "#3e5c8b"),
        axis.text = element_text(colour = "#fffff0"), 
        axis.title = element_text(colour = "#fffff0"),
        plot.title = element_text(colour = "#fffff0", hjust = 0.5),
        panel.grid.major.x = element_line(colour = "#fffff0"),
        panel.grid.major.y = element_line(colour = "fffff0"))

suitable_areas_map 

#ggsave("suitable_areas_formatted.png", plot = suitable_areas_map)

#Use the magick package to trim off the white edges.
library(magick)
suitable_areas_map_trim <- magick::image_read("suitable_areas_formatted.png") %>% magick::image_trim()


#image_write(suitable_areas_map_trim, path = "suitable_areas_map_trim.png", format = "png")

```



```{r}
### save the map ###
#save_map(png_name = "seaweed_gen_suitable.png", title = "General Seaweed Suitability", 
#         color = "light blue")

# sargassum
#save_map(png_name = "sargassum_suitable.png", title = "Sargassum Suitability", 
#         color = "purple")
# Euchema
#save_map(png_name = "euchema_suitable.png", title = "Euchema Suitability", 
#         color = "green")
# Gracilaria
#save_map(png_name = "gracilaria_suitable_L.png", title = "Gracilaria Suitability", 
 #        color = "blue")
```


# Exclusion Layers

```{r}
# Marine protected areas
mpa <- raster(paste0(path,"/mpas_binary_hy.tif"))

# Shipping lanes
shipping_lanes <- raster(paste0(path,"/ship_binary_hy.tif"))

# Submarine Cable Areas
sub_cable_areas <- raster(paste0(path,"/sub_cable_area_binary_hy.tif"))

# Pipeline Areas
pipeline_areas <- raster(paste0(path,"/pipe_binary_hy.tif"))

# Military danger zones
military_danger <- raster(paste0(path,"/danger_binary_hy.tif"))

# Oil and Gas platforms
oil_gas_platforms <- raster(paste0(path,"/og_binary_hy.tif"))

plot(mpa)

plot(shipping_lanes)

plot(military_danger)

```

```{r}
# Marine protected areas exclusion map

# Convert to data frame
mpa_df <- raster::rasterToPoints(mpa) %>%
  as.data.frame() %>% 
  mutate(suit = case_when(
    mpas_binary_hy == 1 ~ "not excluded",
    mpas_binary_hy == 0 ~ "excluded"
  ))


mpa_map <-
  ggplot() +
  geom_raster(data = mpa_df, aes(x = x, y = y, fill = suit)) +
  geom_sf(data = world, fill = "cornsilk3") +
  coord_sf(xlim = c(-100, -80), ylim = c(22, 32)) +
  scale_fill_manual(values = c("rosybrown3", "lightgray")) +
  theme_minimal() +
  labs(fill = "",
       x = "longitude",
       y = "latitude", 
       title = "Marine Protected Areas Exclusion")

mpa_map 

# ggsave("mpa.png", plot = mpa_map)

```

```{r}
# Military Danger Zones

# Convert to data frame
military_df <- raster::rasterToPoints(military_danger) %>%
  as.data.frame() %>% 
  mutate(suit = case_when(
    danger_binary_hy == 1 ~ "not excluded",
    danger_binary_hy == 0 ~ "excluded"
  ))


military_map <-
  ggplot() +
  geom_raster(data = military_df, aes(x = x, y = y, fill = suit)) +
  geom_sf(data = world, fill = "cornsilk3") +
  coord_sf(xlim = c(-100, -80), ylim = c(22, 32)) +
  scale_fill_manual(values = c("rosybrown3", "lightgray")) +
  theme_minimal() +
  labs(fill = "",
       x = "longitude",
       y = "latitude", 
       title = "Military Danger Zones Exclusion")

military_map

#ggsave("military.png", plot = military_map)
  


```


```{r}
# Shipping Lanes

# Convert to data frame
shipping_lanes_df <- raster::rasterToPoints(shipping_lanes) %>%
  as.data.frame() %>% 
  mutate(suit = case_when(
    ship_binary_hy == 1 ~ "not excluded",
    ship_binary_hy == 0 ~ "excluded"
  ))


shipping_lanes_map <-
  ggplot() +
  geom_raster(data = shipping_lanes_df, aes(x = x, y = y, fill = suit)) +
  geom_sf(data = world, fill = "cornsilk3") +
  coord_sf(xlim = c(-100, -80), ylim = c(22, 32)) +
  scale_fill_manual(values = c("rosybrown3", "lightgray")) +
  theme_minimal() +
  labs(fill = "",
       x = "longitude",
       y = "latitude", 
       title = "Shipping Lanes Exclusion")

shipping_lanes_map

#ggsave("shipping_lanes.png", plot = shipping_lanes_map)
  

```


# Eclusion Map

```{r}
exclusion <- overlay( mpa,
                      shipping_lanes,
                      #sub_cable_areas,
                      #pipeline_areas,
                      military_danger,
                      #oil_gas_platforms,
                      fun = function(a, b, c){a*b*c})

#plot(exclusion)
#map('world',fill=T,add=T,col='gray')


# Convert to data frame
exclusion_areas_df <- raster::rasterToPoints(exclusion) %>%
  as.data.frame() %>% 
  mutate(suit = case_when(
    layer == 1 ~ "not excluded",
    layer == 0 ~ "excluded"
  ))


exclusion_areas_map <-
  ggplot() +
  geom_raster(data = exclusion_areas_df, aes(x = x, y = y, fill = suit)) +
  geom_sf(data = world, fill = "cornsilk3") +
  coord_sf(xlim = c(-100, -80), ylim = c(22, 32)) +
  scale_fill_manual(values = c("rosybrown3", "lightgray")) +
  theme_minimal() +
  labs(fill = "",
       x = "longitude",
       y = "latitude", 
       title = "Excluded Seaweed Aquaculture Areas")

exclusion_areas_map 

#ggsave("exclusion_areas.png", plot = exclusion_areas_map)

```

# Final Suitability Map


```{r}
# Get matching resolutions, extents,and columns to overlay suitability and exclusion
np_binary_2 <- resample(np_binary, mpa)
depth_binary_2 <- resample(depth_binary, mpa)
max_sst_binary_2 <- resample(max_sst_binary, mpa)
min_sst_binary_2 <- resample(min_sst_binary, mpa)
sal_min_binary_2 <- resample(sal_min_binary, mpa)
curvel_max_binary_2 <- resample(curvel_max_binary, mpa)
```



```{r}
seaweed_areas <- overlay(np_binary_2, 
                          depth_binary_2, 
                          #max_sst_binary_2,
                          #min_sst_binary_2, 
                          #sal_max_binary, 
                          sal_min_binary_2, 
                          curvel_max_binary_2, 
                          #curvel_min_binary,
                          mpa,
                          shipping_lanes,
                          #sub_cable_areas,
                          #pipeline_areas,
                          military_danger,
                          #oil_gas_platforms,
                          fun = function(a, b, c, d, e, f, g){a*b*c*d*e*f*g})

#plot(suitable_areas)
#map('world',fill=T,add=T,col='gray')

# Convert to data frame
seaweed_df <- raster::rasterToPoints(seaweed_areas) %>%
  as.data.frame() %>% 
  mutate(suit = case_when(
    layer == 0 ~ "unsuitable", TRUE ~ "suitable"
  ))


seaweed_map <-
  ggplot() +
  geom_raster(data = seaweed_df, aes(x = x, y = y, fill = suit)) +
  geom_sf(data = world, fill = "cornsilk3") +
  coord_sf(xlim = c(-100, -80), ylim = c(22, 32)) +
  scale_fill_manual(values = c("olivedrab", "lightgray")) +
  theme_minimal() +
  labs(fill = "",
       x = "longitude",
       y = "latitude", 
       title = "Potential Seaweed Aquaculture Areas")

seaweed_map 

#ggsave("seaweed.png", plot = seaweed_map)

```

```{r}

# Sargassum
#save_map(raster = suitability_zoom,
        # png_name = "sargassum_sui_excl.png", title = "Sargassum Suitability + Exclusion", 
        # color = "purple")

# Gracilaria
#save_map(raster = suitability_zoom, 
         #png_name = "gracilaria_sui_excl_L.png", title = "Gracilaria Full Suitability Map", 
         #color = "blue")

```

